---
title: "Week 12 - Time-Series Cross-Sectional Analysis"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: false
    css: "css/learnr-theme.css"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(learnrhash)
library(texreg)
gradethis:: gradethis_setup()

library(tidyverse)
library(broom)
library(QM2022)
data(df_learnr12_rai)
#load("df_learnr12_rai.rda") # for local testing
```


## Introduction

Welcome to this week's learnr tutorial! You'll practice how to run pooled and fixed effects regression models, compare the results and illustrate them by computing predicted values for a selection of countries.

Before you get started with specifying the models, let's repeat some key concepts of this week.

###

```{r q1, echo = FALSE}
question("Why does it make sense to use a time-series cross-section regression model? Note that more than one answer might be correct.",
  answer("To adress concerns about omitted variable bias", correct = T),
  answer("To account for potential exogenous shocks", correct = T),
  answer("To deal with NAs in our data set"),
  answer("To incorporate unit-specific, time-varying characteristics in our model"),
  random_answer_order = TRUE,
  allow_retry = TRUE,
  incorrect = "Have a look at the lecture material."
)
```

###

```{r q2, echo = FALSE}
question("What is commonly not constant within units over time?",
  answer("The electoral system", 
         message = "How often has the electoral system in your home-country changed in your lifetime?"),
  answer("The unemployment rate", correct = T),
  answer("Culture", 
         message = "Though you may argue that culture is everchanging and subject to the understanding of the individual, for a fixed effects regression we may assume that it is somewhat constant over time."),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Think about how much each answer option usually varies e.g. over the years."
)
```

###

```{r q3, echo = FALSE}
question("When might we want to use a two-way fixed effects model?",
  answer("When we want to test two competing explanations", message = "We can do this with or without a fixed effect."),
  answer("When we want to account for two variables that are constant over time", message = "Two variables that are constant over time, but differ from unit to unit would be covered with a unit-fixed effect."),
  answer("When we want to account for differences between units that are constant over time and for exogenous shocks that affect all units simultaneously", correct = T),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Have a look at the lecture material."
)
```

###

```{r q4, echo = FALSE}
question("What happens when we use a unit-fixed effect?",
  answer("It changes the intercept for each unit", correct = T),
  answer("It changes the slope for each unit"),
  answer("It runs a separate model for each unit"),
  random_answer_order = TRUE,
  allow_retry = T,
  incorrect = "Have a look at the lecture material."
)
```


## Regression models

Góméz-Díaz and Góméz-Díaz (2020) have argued that prolonged membership in the European Union leads to an increase in regional authority in different member states. However, they also include a variable for the length of World Bank membership. In the following, we will focus on the effect of World Bank membership on sub-national authority. 

In the following exercises you will need the following variables:

| Variable name |  Description
|:--------------|:-------------
| `n_RAI` | Regional Authority Index
| `length_EU` |  Length of EU membership in years
| `length_world` | Length of World Bank membership in years
| `democ` | Polity2 Democracy Index
| `gdp_capita` | Gross domestic product per capita
| `population_size` | Size of population
| `nationalization` | Nationalization of the party system
| `country_name` | Name of the country


### Preparing the data and running the first model

The data for the following exercise is stored in an object called `rai`, which is already loaded into your environment. Run a pooled model with `n_RAI` as the dependent variable and `length_world` as the main independent variable of interest. In addition, control for `length_EU`, `democ`, `gdp_capita`, `population_size` and `nationalization`.

```{r exercise1, exercise = TRUE}

```

```{r q5, echo = FALSE}
question_text("What is the coefficient for the length of World Bank membership (`length_world`) in the pooled model? Round the coefficient to two digits after the decimal.",
  answer("-0.04", correct = T),
  allow_retry = TRUE,
  incorrect = "Have you rounded the coefficient correctly?"
)
```

###

Now run the same model, but include a fixed effect for the unit. The units are the countries (`country_name`) in this case. Store the results of the model in an object called `fe_model`.

```{r exercise2, exercise = TRUE}

```


```{r q6, echo = FALSE}
question_text(" What is the unit-specific intercept for Albania? Round the coefficient to two digits after the decimal.",
  answer("0.86", correct = T),
  allow_retry = TRUE,
  incorrect = "Did you correctly specify the fixed effect, making sure there is an intercept for each country?"
)
```

```{r q7, echo = FALSE}
question_text("What is the coefficient for the length of World Bank membership (`length_world`) in the fixed effects model? Round the coefficient to two digits after the decimal.",
  answer("0.03", correct = T),
  allow_retry = TRUE,
  incorrect = "Did you correctly specify the fixed effect, making sure there is an intercept for each country?"
)
```

```{r q8, echo = FALSE}
question_text("If you wanted to present your results in a table using `screenreg()`, but wanted to exclude the country fixed effects from the table, which argument would you use?",
  answer('omit.coef', correct = T),
  answer('omit.coef = "country_name"', correct = T),
  answer('omit.coef="country_name"', correct = T),
  allow_retry = TRUE,
  incorrect = "Take a look at the arguments of the `screenreg()` function."
)
```

```{r q9, echo = FALSE}
question("Comparing the coefficients from the pooled model and the fixed effects model, what do you conclude?",
         answer("Due to omitted variable bias, we would have expected a negative relationship between the length of World Bank membership and regional authority based on the estimate of the pooled model. After accounting for the unobserved, constant variation across countries, we are able to show that the relationship is, in fact, positive and statistically significant.", 
                correct = T),
         answer("Accounting for the unobserved, constant variation across countries has shown that there is no effect of World Bank membership on regional authority.",
                message = "Have a look at the regression results again. Is the effect non-significant?"),
         answer("We have different numbers of observation and therefore the coefficients differ.", 
                message = "Take a look at the number of observations included in each model. Are they the same or do they differ?"),
         answer("The coefficients are the same.", message = "If the coefficients are the same, you may want to retrace your steps and check your code whether there is an error."),
         random_answer_order = TRUE,
         allow_retry = T
)
```


## Prediction

Remember, predicted values can help us to communicate the results of our analyses. In the following section, you have to calculate predicted values and confidence intervals for the maximum length of World Bank membership for Uruguay and Japan, holding everything else constant. 

*You can re-run the regression model in the chunk below. Alternatively, the correct regression output is already stored in an object called `fe_model`.*

```{r exercise3-setup}
fe_model <-  lm(n_RAI ~ length_world + length_EU + democ + gdp_capita + 
                  population_size + nationalization + country_name-1, 
                data = rai)
```

```{r exercise3, exercise = TRUE}

```

```{r q10, echo = FALSE}
question("Compare the predicted values. Which country has a higher predicted value for sub-national authority?",
  answer("Japan", 
         message = "Keep in mind that you also have to take the uncertainty around the predicted value into account before drawing conclusion. Have a look at the confidence intervals."),
  answer("Uruguay", 
         message = "Keep in mind that you also have to take the uncertainty around the predicted value into account before drawing conclusion. Have a look at the confidence intervals."),
  answer("Unclear, because the confidence intervals overlap.", 
         correct = T),
  random_answer_order = TRUE,
  allow_retry = T)
```

###

Now, calculate predicted values for regional authority for the minimum and maximum values of length of World Bank membership for Japan, holding everything else constant.

*You can re-run the regression model in the chunk below. Alternatively, the correct regression output is already stored in an object called `fe_model`.*

```{r exercise4-setup}
fe_model <-  lm(n_RAI ~ length_world + length_EU + democ + gdp_capita + 
                  population_size + nationalization + country_name-1, 
                data = rai)
```

```{r exercise4, exercise = TRUE}


```

```{r q11, echo = FALSE}
question_text("What is the difference in predicted values for minimum and maximum length of World Bank membership for Japan? Round to two digits after the decimal",
  answer("2.10", correct = T),
  answer("2.1", correct = T),
  allow_retry = T,
  incorrect = "Did you correctly specify the scenarios?")
```


## Submit

```{r context="server"}
learnrhash::encoder_logic(strip_output = TRUE)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(default_ui(url = "https://ilias.uni-mannheim.de/goto.php?target=svy_1245812&client_id=ILIAS"))
```